**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**

### 2025-08-19 paper/manuscript.qmd
final

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Emma Benn 
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Maisie Bennett
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  Interoception, the sense of the body's internal state, is crucial for mental and physical health, yet its measurement is hindered by tools with conceptual and psychometric issues. To address these limitations, we developed and validated the Multimodal Interoception Questionnaire (Mint). In Study 1 (N=559), we generated a comprehensive item pool using a novel "modality-by-context" framework and employed the recently developed psychometric network analysis for data-driven item reduction. In Study 2 (N=737), we validated the resulting instrument in an independent sample, establishing its structure and comparing its predictive power against established interoception questionnaires (MAIA-2, BPQ, IAS) across a wide range of clinical and cognitive outcomes. This rigorous process yielded a 33-item scale with a stable hierarchical structure comprising three metaclusters (Interoceptive Deficit, Interoceptive Awareness, Visceroception) and 11 distinct facets. The Mint demonstrated superior criterion validity, consistently outperforming existing scales in predicting outcomes tied to altered bodily perception, including alexithymia, ADHD, autism, and somatic symptom clusters. The Mint offers a psychometrically robust and nuanced measure of self-reported interoception that integrates and extends existing scales, providing researchers and clinicians with a practical and readily usable tool to investigate the role of bodily sensations in health and disease.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness, Health"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

@@

Interoception - the sensing and interpretation of the body's internal physiological signals - plays a fundamental role in processes ranging from homeostatic regulation to emotion, decision-making, and mental health [@craig2002how; @critchley2017interoception]. Despite its growing recognition, the field is hindered by inconsistent definitions and fragmented measurement approaches [@khalsa2018interoception; @desmedt2025discrepancies; @desmedt2022measures]. Available tools include "objective" physiological tasks (e.g., heartbeat counting or tracking) and "subjective" self-reports, and encompass both "explicit" paradigms that direct participants' attention and consciousness to internal sensations and "implicit" paradigms that do not (e.g., heartbeat evoked potentials recorded during rest). These assessments are often targeted at only one of several interoceptive modalities (e.g., cardioception, respiroception, gastroception), and are mapped onto theoretical dimensions [e.g., accuracy, sensitivity, awareness, @garfinkel2015knowing] within a rapidly evolving theoretical scaffolding. However, no consensus exists on a gold-standard measure, and each approach has distinct limitations: physiological tasks often require specialised equipment and expertise, are time-consuming, and are challenging to administer at scale, whereas questionnaires are constrained by their inherent subjectivity. Crucially, the psychometric quality of many of these measures is questionable [e.g., see criticism about the heartbeat counting task, @zamariola2018interoceptive], yielding low inter-task correlations [@murphy2018classifying], raising serious concerns about their reliability and validity (e.g., measures presented as measuring one theoretical dimension but likely measuring something else). This measurement conundrum hinders progress in comparing, replicating, and interpreting interoception-related findings, underscoring the need to develop improved tools that align with modern psychometric standards.

<!-- There are many questionnaire measures – 14 of which are commonly used and 5 (BPQ, MAIA, BAQ, SAQ, and PBCS) which cover 86.5% of all citations of interoceptive sensibility questionnaires – Desmedt et al., 2022 -->
The present work focuses on self-report questionnaires, which offer a scalable and accessible tool complementing physiological tasks. Among the numerous questionnaires available, a handful seem to dominate the literature [@desmedt2022measures], including the Body Perception Questionnaire [BPQ, @kolacz2018body], the Multidimensional Assessment of Interoceptive Awareness [MAIA, @mehling2018multidimensional], and the Interoceptive Accuracy Scale [IAS, @murphy2018classifying]. These tools were developed in different eras of a rapidly evolving field, each with its own objectives and philosophies. The BPQ, one of the earliest measures, emphasizes the assessment of body awareness and autonomic nervous system (ANS) reactivity, focusing on subjective experiences of physiological stress reactions in organs innervated by the ANS, such as those triggered by stress or trauma [@kolacz2018body]. The MAIA incorporated ideas from the then-booming emotion regulation and mindfulness fields, including subscales assessing adaptive coping strategies and metacognitive insights about the body. By contrast, the IAS was recently crafted to sidestep such high-level - possibly tangential - constructs and instead capture subjective reports of concrete bodily events (e.g., coughing, sneezing) as proxies for interoceptive accuracy.

These differing philosophies have practical consequences, leading to poor convergence between scales that are assumed to measure the same construct [@vig2022questionnaires] - a classic example of the "jingle fallacy" where the shared label of "interoception" can mask underlying measurand differences, which in turn hinders replication and theoretical integration. Recent work has been focused on aligning subjective measures and facets with theoretical models, debating whether they measure interoceptive "sensitivity" or other dimensions (accuracy, attention, deficits). However, this focus on a theory-driven approach might be premature when theoretical models are in flux, and when a topographic mapping to self-report dimensions is not guaranteed (e.g. distinctions that are relevant at a neuroanatomical or cognitive level might not necessarily translate into the same structural landscape of subjective experiences). More fundamentally, existing questionnaires suffer from two major structural blind spots. First, they are often narrow or unsystematic in their coverage of bodily *modalities*, with a heavy bias towards cardio-respiratory sensations and a neglect of others like gastric, urogenital, or thermoregulatory signals. Second, and perhaps most critically, they fail to control for the *context* in which an interoceptive experience occurs, introducing significant variability in how participants interpret and respond to items. For example, an item about "feeling your heart beat" is ambiguous; its frequency, experience and its interpretation differ vastly depending on whether it occurs during physical exercise, a moment of anxiety, or quiet rest (e.g., someone who exercises regularly, or is anxious, might more easily recall an instance of occurrence of feeling their heart beating and respond based on that contextual saliency).

To address these gaps, we introduce the Multimodal Interoception Questionnaire (Mint). The overarching aim of this work is to develop and validate a new questionnaire that systematically addresses these limitations. In Study 1, we detail the bottom-up development of the Mint, using a "modality-by-context" framework to generate a comprehensive item pool, which is then refined using cutting-edge psychometric network analysis. In Study 2, we validate the final version of the Mint against the established interoception questionnaires, demonstrating its robust structure and predictive validity for a range of clinically-relevant outcomes.

@@

The primary goal of Study 1 was to develop a new interoception questionnaire by generating a comprehensive item pool and refining it into a psychometrically robust scale. To overcome the limitations of existing measures, we adopted a bottom-up, theoretically agnostic approach. Instead of imposing a pre-defined theoretical structure, we focused on two tangible and systematically controllable aspects of interoceptive experience: the *modality* of the sensation and the *context* in which it occurs. The objective was not to create *yet another scale* measuring an alleged distinct construct or variation, but rather to synthesize existing ideas and frameworks into a comprehensive and modern tool that captures the richness of interoceptive experiences and encompasses the most relevant facets of interoception.

<!-- Give examples of items for each part of the framework here so that the method makes more sense? -->
Our item generation process was guided by a "modality-by-context" grid designed to overcome the limitations of existing scales. For *modalities*, we expanded beyond the typical cardio-respiratory focus to include a broader range of bodily systems: cardiac, respiratory, gastric, genital, bladder & colon, and skin & temperature [the latter of which being increasingly recognized as a key interoceptive channel, @crucianelli2023role]. We also included a "general state" category to capture the integrated sense of the body's overall physiological condition [@craig2002how; @craig2003interoception]. For *contexts*, we designed facets to capture specific situational contexts like negative (anxious) and positive (sexual) arousal, which reflect distinct bodily states with differing interoceptive signatures, as well as other experiential qualities tapping into elaboration and threshold of perception, such as sensitivity, accuracy, and confusion. To maintain continuity with established work while making targeted improvements, we purposefully adapted items from existing questionnaires (e.g., IAS-like items related to observable behaviors). This systematic framework yielded a diverse pool of 120 initial items, ensuring comprehensive coverage while minimizing the ambiguity inherent in context-free questions.

The final step involved a data-driven reduction of this item pool. We employed cutting-edge psychometric network-based techniques to identify and remove redundant items and to uncover the most stable and coherent dimensional structure. Additionally, we designed the study so that we could explore an often overlooked psychometric quality: the invariance of the structure to the presentation format (how items are grouped together on the screen, whether by modality, facet, or randomly) influences the resulting psychometric structure. This rigorous, multi-stage process was designed to produce a subset of items with a clear, stable, and theoretically rich hierarchical structure, ready for formal validation.

@@

<!-- summary -->
The primary aim of Study 1 was to develop and refine a large, systematically generated pool of interoceptive items into a psychometrically sound and structurally coherent questionnaire. Starting with 120 items derived from a modality-by-context grid, we employed a multi-step, data-driven reduction process using cutting-edge structure analysis methods. This rigorous procedure yielded a final, balanced set of 45 items, organized within a stable hierarchical structure comprising three higher-order dimensions (Interoceptive Deficits, Interoceptive Awareness, Interoceptive Sensitivity) and 12 distinct lower-level facets.

<!-- Effect of item presentation -->
One innovation of this study was to test the effect of item presentation by comparing structures across different grouping conditions. While the analysis revealed no significant effect, a consistent (though non-significant) trend suggested that grouping items by modality may be slightly more conducive to revealing the underlying psychometric structure than grouping by facet or randomly. This finding, while preliminary due to the sample size per condition ($N \pm 200$), highlights a potentially important and under-investigated aspect of questionnaire design that warrants future research.
<!-- The final reduction from 54 to 45 items was a consensus-based decision to remove three clusters for reasons of conceptual clarity and parsimony, a necessary step to ensure the final scale's usability and content validity. -->

The resulting hierarchical structure offers novel yet interpretable insights into self-reported interoception, with flexible degrees of granularity. Notably, we categorised several facets (typically related to objective external behaviours, like coughing or sneezing) under the label "Awareness", that might traditionally be named "Accuracy" in other scales. Our rationale is that subjective self-report, by its very nature, cannot be a pure measure of objective accuracy, but at best a proxy of an individual's subjective certainty or awareness of their interoceptive processes. Furthermore, our bottom-up approach revealed intriguing facets, such as "Olfactory Compensation". This finding provide further evidence for a pseudo-interoceptive pathway, whereby individuals with potentially poorer access to internal signals may compensate by relying on external cues (e.g., odours) to infer their bodily state. The "Sensitivity" metacluster also grouped core visceroceptive facets (e.g., "Cardioception", "Gastroception") with "Dermal Hypersensitivity", pointing to a potential common dimension of heightened perceptual sensitivity to bodily sensations.

In summary, Study 1 successfully leveraged a systematic item generation process and modern analytical techniques to produce a 45-item questionnaire with a clear, stable, and theoretically rich hierarchical structure. The next crucial step is to formally validate this new instrument by examining its psychometric properties and its relationship with existing measures in a new, independent sample, which is the objective of Study 2.

@@

Following the generation and selection of a set of 45 interoceptive items, the next objective is to formally validate this new instrument in a large, independent sample, a crucial step for establishing the psychometric soundness and unique contribution of the Mint scale to the field of interoception research. To this end, we sought to 1) re-assess the structure of the shorter version and 2) establish its convergent validity by examining the correlations with other established and popular interoception questionnaires (namely the MAIA-2, BPQ, and IAS), as well as other related measures (e.g., alexithymia). Finally, we assessed its criterion validity by directly comparing the predictive power of the interoceptive scales across a broad range of outcomes, including emotional, cognitive, and clinical variables.

@@

The primary goal of Study 2 was to validate the items selected in Study 1. This involved re-assessing its structure in a new, large sample, and establishing its convergent and criterion validity against other established interoception questionnaires. The analysis led to a final, refined 33-item Mint scale with a stable and theoretically coherent hierarchical structure, which demonstrated substantial predictive power across a range of outcomes.

Re-validating the structure after item selection is a crucial step in questionnaire development. When a large item pool is reduced, the context changes, which can alter the perceived meaning and relationships between the remaining items. For instance, a questionnaire with many items suggesting abnormality or pathology might colour other neutral items with a negative connotation. While the original structure was mostly replicated, the re-validation allowed for a clarification of some of the higher-order dimensions (for instance, the "Sensitivity" metacluster became more clearly defined around core interoceptive clusters). The final structure, comprising 11 facets and 3 metaclusters, showed excellent structural consistency and stability, providing a solid foundation for the scale. However, the *Olfa* (Olfactory Compensation) cluster showed lower stability in its placement within the *Deficit* metacluster. This suggests its role may be more complex, and future studies are needed to investigate its place and invariance across different populations and contexts.

The criterion validity analysis revealed a clear and compelling pattern of results. The Mint scale, particularly its *Deficit* metacluster, consistently outperformed other scales in predicting outcomes directly related to alterations in bodily experience and self-regulation, such as alexithymia, dissociative experiences, ADHD, autism, and various somatic symptoms. The main contender, the MAIA-2, showed an advantage primarily for outcomes related to general well-being, such as life satisfaction, depression, and anxiety. However, this advantage was almost entirely driven by dimensions highly related to emotion regulation abilities, namely *Trusting* and *Not-Worrying*, whose inclusion as core interoceptive facets has been questioned [@ferentzi2021examining]. When these dimensions were removed (in the *iMAIA* version), the MAIA-2's predictive power for these outcomes dropped considerably, often below that of the Mint.

This highlights a crucial issue in the field: the potential for misattributing findings to "interoception" when they may be driven by broader emotion regulation capacities [@murphy2020testing]. While valuable, the MAIA-2's most predictive subscales appear less related to core interoceptive processing and more to metacognitive beliefs about one's body and emotional coping. The Mint, in contrast, appears to tap more directly into distinct facets of interoceptive experience, from visceral sensitivity to perceived deficits. The pattern is clear: for high-level, metacognitive aspects of affective state and well-being, the MAIA-2's emotion regulation facets are highly relevant. For outcomes tied more specifically to bodily processes, clinical conditions characterized by sensory dysregulation, and nuanced interoceptive experiences, the Mint offers superior predictive validity and specificity.

@@

Across two studies, we developed and validated the Multimodal Interoception Questionnaire (Mint), a new 33-item self-report measure of interoception. Our approach addressed key limitations of existing scales by starting with a systematically generated, broad item pool that explicitly controlled for context and modality, and then using a data-driven reduction process. The result is a psychometrically robust instrument with a stable, hierarchical structure comprising three higher-order metaclusters, namely *Interoceptive Deficit*, *Interoceptive Awareness*, and *Visceroception*, and 11 distinct lower-level facets. The validation study demonstrated that the Mint not only captures variance present in other popular scales (MAIA-2, BPQ and IAS) but also offers superior predictive power for a range of clinical and cognitive outcomes, particularly those tied to altered bodily perception.

A key contribution of the Mint is its potential ability to disentangle core interoceptive experiences from more general emotion regulation capacities. Our findings show that while the MAIA-2 is a potent predictor of general well-being, this is largely driven by its *Trusting* and *Not-Worrying* subscales, which heavily tap into emotion regulation skills and coping strategies. The Mint, conversely, excelled at predicting conditions like alexithymia, ADHD, autism, and somatic symptoms, where alterations in bodily sensing are a core feature. This suggests the Mint provides a more specific measure of interoceptive processing, offering researchers a finer-grained tool to probe the role of bodily awareness in health and disease.

Despite these strengths, the present work has limitations that open avenues for future research. First, as a self-report measure, the Mint captures subjective beliefs about interoceptive abilities, not objective accuracy. A crucial next step is to investigate the relationship between Mint scores and physiological or behavioural measures of interoception (e.g., heartbeat tracking tasks, heartbeat evoked potentials). This would help establish the extent to which subjective reports on the Mint correspond to objective interoceptive performance. Future research could also systematically examine the cost of shorter versions, by investigating the tradeoffs incurred by versions with only the one or two most central facets per metacluster and their impact on psychometric qualities and predictive power.

>>>>

---
title: "**The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS**"
shorttitle: "Mint Validation"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Investigation
      - Methodology
      - Project administration
      - Supervision
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Emma Benn 
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Maisie Bennett
    role:
      - Investigation
    affiliations:
      - ref: "id1"
  - name: Giulia Poreiro
    orcid: 0000-0002-2343-5109
    role:
      - Investigation
      - Methodology
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  Interoception, the sensing of the body's internal state, is an important component of homeostatic and allostatic regulation with wide-ranging implications for mental and physical health, yet its measurement is hindered by tools with conceptual and psychometric issues. To address these limitations, we developed and validated the Multimodal Interoception Questionnaire (Mint). In Study 1 (N=559), a comprehensive item pool using a novel "modality-by-context" framework was generated and then reduced using psychometric network analysis. In Study 2 (N=737), we validated the Mint in an independent sample, establishing its structure,  comparing its predictive power against established interoception questionnaires (MAIA-2, BPQ, IAS), and across a wide range of clinical and cognitive outcomes. This rigorous process yielded a 33-item scale with a stable hierarchical structure comprising three metaclusters (Interoceptive Deficit, Interoceptive Awareness, Visceroception) and 11 distinct facets. The Mint demonstrated superior criterion validity, consistently outperforming existing scales in predicting outcomes tied to altered bodily perception, including alexithymia, ADHD, autism, and somatic symptom clusters. The Mint offers a psychometrically robust and nuanced measure of self-reported interoception that integrates and extends existing scales, providing researchers and clinicians with a practical and readily usable tool to investigate the role of bodily sensations in health and disease.
#   **Significance Statement**. TODO.
keywords: "Interoception questionnaire, interoceptive accuracy scale, MAIA, Mint Validation, Body Awareness, Health"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

@@

Interoception - the sensing and interpretation of the body's internal physiological signals - plays a fundamental role in processes ranging from homeostatic regulation to emotion, decision-making, and mental health [@craig2002how; @critchley2017interoception]. Despite its growing recognition, the field has been hindered by inconsistent definitions and fragmented measurement approaches [@khalsa2018interoception; @desmedt2025discrepancies; @desmedt2022measures]. Available tools include "objective" physiological tasks (e.g., heartbeat counting or tracking) and "subjective" self-reports, and encompass both "explicit" paradigms that direct participants' attention to internal sensations and "implicit" paradigms that do not (e.g., heartbeat evoked potentials recorded during rest). These assessments are often targeted at only one of several interoceptive modalities (e.g., cardioception, respiroception, gastroception), and are mapped onto theoretical dimensions [e.g., accuracy, sensitivity, awareness, @garfinkel2015knowing] within a rapidly evolving theoretical landscape. However, no consensus exists on a gold-standard measure witheach approach having distinct limitations. For example, physiological tasks often require specialised equipment and expertise, are time-consuming, and are challenging to administer at scale, whereas questionnaires are constrained by their inherent subjectivity. Crucially, the psychometric quality of many of these measures is questionable [e.g., see criticism about the heartbeat counting task, @zamariola2018interoceptive], yielding low inter-task correlations [@murphy2018classifying] and raising important concerns regarding their reliability and validity (e.g., measures presented as measuring one theoretical dimension but likely measuring something else). This lack of conceptual clarity and measurement convergence impedes scientific progress in comparing, replicating, and interpreting interoception-related findings, emphasising the need to develop improved tools that align with modern psychometric standards.

<!-- There are many questionnaire measures – 14 of which are commonly used and 5 (BPQ, MAIA, BAQ, SAQ, and PBCS) which cover 86.5% of all citations of interoceptive sensibility questionnaires – Desmedt et al., 2022 -->
The present work focuses on self-report questionnaires, which offer a scalable and accessible tool complementing physiological tasks. Among the numerous interoceptive questionnaires available, a handful dominate the literature [@desmedt2022measures], including the Body Perception Questionnaire [BPQ, @kolacz2018body], the Multidimensional Assessment of Interoceptive Awareness [MAIA, @mehling2018multidimensional], and the Interoceptive Accuracy Scale [IAS, @murphy2018classifying]. These tools were developed in different eras of a rapidly evolving field, each with its own objectives and philosophies. For example, the BPQ, one of the earliest measures, captures bodily awareness and autonomic nervous system (ANS) reactivity by focusing on experiences of physiological stress reactions in organs innervated by the ANS, such as those triggered by stress or trauma [@kolacz2018body]. The MAIA incorporated ideas from the then-booming emotion regulation and mindfulness fields, including subscales assessing adaptive coping strategies and metacognitive insights about the body. In contrast, the IAS was recently developed to circumvent these high-level - possibly tangential - constructs, instead capturing subjective reports of concrete bodily events (e.g., coughing, sneezing) as proxies for interoceptive accuracy.

These differing philosophies have practical consequences, leading to poor convergence between scales that are assumed to measure the same construct [@vig2022questionnaires] - a classic example of the "jingle fallacy" where the shared label of "interoception" can mask underlying measurand differences, hindering replication and theoretical integration. In acknowledgement of this issue, efforts have been made to align subjective measures with theoretical models that outline different facets of interoception (e.g., accuracy, attention, deficits). However, focusing on developing measures that are theory-driven may however be premature when theoretical models are in flux, and when a topographic mapping to self-report dimensions is not guaranteed. For example,  interoceptive distinctions that are relevant at a neuroanatomical or cognitive level might not necessarily map onto the structure of subjective experiences. More fundamentally, existing interoception questionnaires typically fail to consider the full range of interoceptive modalities and the context in which interoceptive experiences occur. Most existing questionnaires focus on a narrow range of bodily modalities, with a heavy bias towards cardio-respiratory sensations at the expense of others like gastric, urogenital, or thermoregulatory signals. Perhaps most critically, existing measures ask participants to make general assessments across most/all contexts, which does not capture the potential homeostatic relevance of attending or detecting internal signals. Without a contextual reference point, there is likely to be substantial variability in how participants interpret and respond to items. For example, a general item about "feeling your heart beat" is ambiguous since its salience, frequency, experience and interpretation would differ depending on the context, e.g., on whether it occurs during physical exercise, a moment of anxiety, or quiet rest (e.g., someone who exercises regularly, or is anxious, might more easily recall an instance of occurrence of feeling their heart beating and respond based on that contextual saliency).

To address these gaps, we developed and validated the Multimodal Interoception Questionnaire (Mint), a new questionnaire that systematically addresses the above limitations. In Study 1, we describe the bottom-up development of the Mint, using a "modality-by-context" framework to generate a comprehensive item pool, which was then refined using cutting-edge psychometric network analysis. In Study 2, we validated the final 33-item version of the Mint against the established interoception questionnaires, demonstrating its robust structure and predictive validity for a range of clinically-relevant outcomes.

@@

The aim of Study 1 was to develop a new interoception questionnaire by generating a comprehensive item pool and subsequently refining it into a psychometrically robust scale. To overcome the limitations of existing questionnaires, we adopted a bottom-up, theoretically agnostic approach. Rather than imposing a pre-defined theoretical structure on questionnaire items, we instead focused on two tangible and systematically controllable aspects of interoceptive experience: the *modality* of the sensation and the *context* in which it occurs. Our objective was not to develop a questionnaire measuring a new or theoretically distinct facet, but instead build on existing research and synthesise existing ideas and frameworks into a comprehensive and modern tool that captures both the richness of interoceptive experiences and the most prominent facets of interoception.

Item generation was guided by a "modality-by-context" grid. For *modalities*, we expanded beyond the typical cardio-respiratory focus to encompass a wide range of bodily systems: cardiac, respiratory, gastric, genital, bladder & colon, and skin & temperature [the latter of which being increasingly recognized as a key interoceptive channel, @crucianelli2023role]. We also included a "general state" category to capture the integrated sense of the body's overall physiological condition [@craig2002how; @craig2003interoception]. For *contexts*, we designed facets to capture specific situational contexts like negative (anxious) and positive (sexual) arousal, which reflect distinct bodily states with differing interoceptive signatures. Facets were also designed to capture other experiential qualities tapping into elaboration and threshold of perception, such as sensitivity, accuracy, and confusion. This systematic framework yielded a diverse pool of 120 initial items, ensuring comprehensive coverage while minimising the ambiguity inherent in context-free questions.

The initial pool of 120 was then reduced using cutting-edge psychometric network-based techniques, identifying and removing redundant items and revealing  the most stable and coherent dimensional structure. We also designed the study to enable an examination of structural invariance depending on presentation format (e.g., how items are grouped together, whether by modality, facet or randomly) which is an often overlooked psychometric quality. This rigorous, multi-stage process was designed to produce a subset of items with a clear, stable, and theoretically rich hierarchical structure, ready for formal validation in Study 2.

@@

<!-- summary -->
The aim of Study 1 was to develop and refine a large, systematically generated pool of interoceptive items into a psychometrically sound and structurally coherent questionnaire. Starting with 120 items derived from a modality-by-context grid, we employed a multi-step, data-driven reduction process using cutting-edge structure analysis methods. This procedure yielded a final, balanced set of 45 items, organised within a stable hierarchical structure comprising three higher-order dimensions (Interoceptive Deficits, Interoceptive Awareness, Interoceptive Sensitivity) and 12 distinct lower-level facets.

<!-- The final reduction from 54 to 45 items was a consensus-based decision to remove three clusters for reasons of conceptual clarity and parsimony, a necessary step to ensure the final scale's usability and content validity. -->

The resulting hierarchical structure offers novel yet interpretable insights into self-reported interoception, with flexible degrees of granularity. Notably, we categorised several facets (typically related to objective external behaviours, like coughing or sneezing) under the label "Awareness". Although other questionnaires might refer to this as "Accuracy", we preferred to preserve the notion that subjective self-reports, by their very nature, cannot be a pure measure of objective accuracy and are, at best, a proxy for subjective certainty or awareness of interoceptive processes. Furthermore, our bottom-up approach revealed intriguing facets, such as "Olfactory Compensation". This finding provides evidence for a pseudo-interoceptive pathway, whereby individuals with poorer access to internal signals may compensate by relying on external cues (e.g., odours) to infer their bodily state. The "Sensitivity" metacluster also grouped core visceroceptive facets (e.g., "Cardioception", "Gastroception") with "Dermal Hypersensitivity", pointing to a potential common dimension of heightened perceptual sensitivity to bodily sensations.

<!-- Effect of item presentation -->
An additional innovation of Study 1 was examining the effect of item presentation on the resulting psychometric structure by comparing structures across different grouping conditions. Although there were no significant effects of item presentation, a consistent (though non-significant) trend suggested that grouping items by modality (rather than by facet or randomly) could be more conducive to revealing the underlying psychometric structure. This finding, while preliminary due to the sample size per condition ($N \pm 200$), highlights a potentially important and under-investigated aspect of questionnaire design worthy of future research.

@@

Study 1 successfully leveraged a systematic item generation process and modern analytical techniques to produce a 45-item interoception questionnaire with a clear, stable, and theoretically rich hierarchical structure. Study 2 then sought to formally validate the Mint scale by examining its psychometric properties and its relationship with existing measures in a large, independent sample. To this end, we sought to 1) re-assess the structure of the shorter Mint version and 2) establish its convergent validity by examining correlations with other established and popular interoception questionnaires (namely the MAIA-2, BPQ, and IAS), as well as other related measures (e.g., alexithymia). Finally, we assessed the Mint’s criterion validity by directly comparing the predictive power of all interoception questionnaires across a broad range of outcomes, including emotional, cognitive, and clinical variables.

@@

Study 2 sought to formally validate the Mint in a large, independent sample, a crucial step for establishing the psychometric soundness and unique contribution of the scale for interoception research. This involved re-assessing its structure in a new sample and establishing its convergent and criterion validity against other existing interoception questionnaires. The analysis led to a final, refined 33-item Mint scale with a stable and theoretically coherent hierarchical structure, which demonstrated substantial predictive power across a range of outcomes.

Structure re-validation is an important step following item reduction since reducing a large item pool can alter the perceived meaning and relationships between the remaining items. For instance, a questionnaire with many items suggesting abnormality or pathology might influence the interpretation of otherwise neutral items. While the original structure of the Mint was mostly replicated, the re-validation facilitated the clarification of several higher-order dimensions (e.g., the "Sensitivity" metacluster became more clearly defined around core interoceptive clusters). The final Mint structure, comprising 11 facets within 3 metaclusters, yielded excellent structural consistency and stability. The *Olfa* (Olfactory Compensation) cluster had lower stability within the *Deficit* metacluster, possibly indicative of a more complex role. For instance, this cluster could be related to adaptive compensatory processes in some participants and maladaptive strategies in others. Future studies should further investigate its place and invariance across different populations and contexts.

The criterion validity analysis revealed a clear and compelling pattern of results. The Mint scale, particularly its *Deficit* metacluster, consistently outperformed existing interoceptive scales in predicting outcomes directly related to alterations in bodily experience and self-regulation, including alexithymia, dissociative experiences, ADHD, autism, and various somatic symptoms. The MAIA-2 showed an advantage over the Mint for predicting outcomes related to general well-being (e.g., life satisfaction, depression, and anxiety). However, its predictive superiority was almost entirely driven by dimensions related to emotion regulation abilities, namely *Trusting* and *Not-Worrying*, whose inclusion as core interoceptive facets have been questioned [@ferentzi2021examining]. While valuable, the MAIA-2's most predictive subscales appear less related to core interoceptive processing and more to metacognitive beliefs about one's body and emotional coping. The Mint, in contrast, appears to tap more directly into distinct facets of interoceptive experience, from visceral sensitivity to perceived deficits. Indeed, when the emotion-regulation dimensions were removed (in the *iMAIA* version), the MAIA-2's predictive power for well-being outcomes dropped considerably, often below that of the Mint. Overall, for high-level, metacognitive aspects of affective state and well-being, the MAIA-2's emotion regulation facets are highly informative. For outcomes tied more specifically to bodily processes, clinical conditions characterised by sensory dysregulation, and nuanced interoceptive experiences, the Mint offers superior predictive validity and specificity.

@@

Across two studies, we developed and validated the Multimodal Interoception Questionnaire (Mint), a new 33-item self-report measure of interoception. Our approach addressed key limitations of existing questionnaires by starting with a systematically generated, broad item pool that explicitly accounted for context and assessed a wide range of interoceptive modalities.  The Mint is a psychometrically robust instrument with a stable, hierarchical structure comprising three higher-order metaclusters, namely *Interoceptive Deficit*, *Interoceptive Awareness*, and *Visceroception*, and 11 distinct lower-level facets. Study 2 validated the Mint, demonstrating that it not only captures variance present in other popular scales (MAIA-2, BPQ and IAS) but also offers superior predictive power for a range of clinical and cognitive outcomes, particularly those tied to altered bodily perception.

A key advantage of the Mint is its potential ability to disentangle core interoceptive experiences from more general emotion regulation abilities, an identified problem in interoception research where findings may be misattributed to "interoception" when they are driven by emotion regulation capacities [@murphy2020testing]. Our findings show that although the MAIA-2 is a strong predictor of general well-being, this predictive power is driven by its *Trusting* and *Not-Worrying* subscales, which heavily tap into emotion regulation skills and coping strategies. The Mint, in contrast, taps more directly into interoceptive components, and has excellent predictive power for conditions where alterations in bodily sensing are a core feature, including alexithymia, ADHD, autism, and somatic symptoms. Overall,  the Mint provides a more specific measure of self-reported interoceptive processing across contexts and modalities, offering researchers a granular and comprehensive tool with which to further understand the contribution of bodily awareness to human health and disease.

We note several limitations that could serve as the impetus for future research efforts. First, as a self-report measure, the Mint captures subjective beliefs about interoceptive abilities, not objective accuracy. A natural next step will be to establish the extent to which subjective reports on the Mint correspond to objective interoceptive performance (e.g., heartbeat tracking tasks, heartbeat evoked potentials). Second, future research could systematically examine the usefulness of shorter versions of the Mint, for example, by investigating the tradeoffs incurred by versions with only the one or two most central facets per metacluster on psychometric qualities and predictive power.
